Speculation in foreign exchange: noise or information

### 2015-10-13 Knitr/NewPaper.Rnw
Corrected paper.  Now need to write conclusion

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is part of the price-discovery process rather than noise.      
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices. 
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.

Towards the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study:  Extreme (High $99^{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is at or above the $99^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES1}
\end{figure}

This paper contributes to the debate about the nature of speculation in a number of ways. It provides clear evidence that speculation is more informed than uninformed. A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.  Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk. However, extremes are followed by an increase in the variance of exchange rate returns. This suggests that information from informed speculation contributes to the process of price discovery.

Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme. An event study is conducted with the speculative extreme being the event $(t_0)$, sentiment is measured by the price of currency call options relative to the equivalent put.  The foreign exchange rate has a contemporaneous movement with the sentiment.  This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief. However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.  As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.

%In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation is central to many contemporary questions about international financial markets. For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.  If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; uninformed speculators will increase the tendency towards activity that is focused on the short-term.  However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the process of price discovery. In this case, increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any benefits.

@@

\section{Informed and uninformed speculation}
Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the mechanism for the operation of rational expectations. However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making. A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.   Speculation and trading can be categorised as being informed or uninformed. Those trading without information are also called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. 
%these two paragraphs above may go later and just a very short coverage of the distinction between noise-traders and liquidity traders. 
The distinction between informed and uninformed traders is a central component of microstructure research.  The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.  This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.  In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.

@@

\citet[p.529]{BlackNoise} suggested that this interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.  Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{With market-makers using Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

@@

Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.  Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.   This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred from the buying and selling conducted by other traders.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}

Liquidity trading and institutional constraints may mean that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.

@@

The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators. The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information. The aim of this paper is to identify the Friedman-effect.

\section{Measuring speculation and the event study}
The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed. Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.  If informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium; the following price performance should be a random walk.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.

The \citet[pp. 14-15]{Delong1990noise} model is summarised by Equation \ref{eq:NP2}. There are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.  In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put. The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

@@

\subsection{Measuring the proportion of speculators}
Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.  However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.  The Commodity Futures Trading Commission (CFTC) collect weekly information about the open positions held by private entities in the US derivative markets.   A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).\footnote{The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).  Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.  These could be seen as natural hedgers.  Non-commercial accounts are generally considered to be speculators.  See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  The Euro data runs from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}

\section{Analysis of the results}
An \emph{event study} is used to assess the effect of extreme speculation on future returns.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study. The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.  Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.

Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the next two are the cumulative average returns for the whole window (CARW) and the cumulative average returns for the period of the event and afterwards (CARA).  The windows were chosen to correspond to one day and one week of trading.  However, Figure \ref{fig:ES1} makes it clear that altering the size of the event window does not change the findings to any significant degree.

As speculators become more extreme in their opinion, exchange rates are driven in the direction of sentiment. Looking down the seventh column (the 4-day CARW) shows the average cumulative return for the period from four days before the extreme event to the four days after the extreme for extreme sentiment highs and lows of each currency pair.   The extreme highs are (in this case) those points when the risk-reversal skew is at or above the $95^{th}$ percentile.  For the whole event window (CARW), extreme highs (labelled "Hi") are associated with positive returns and the extreme lows (labelled "Lo") are associated with negative returns. In most cases these positive and negative returns appear to be statistically significant. Significance is identified as those cases where more than 95\% of 1000 random bootstrap samples taken from the events being above zero. Mean values of the cumulative returns are calculated from samples of the event days and the distribution of these means are used to construct the confidence intervals.  Returns that are statistically different from zero are identified by an asterisk.

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is informed and part of the process of price-discovery rather than uninformed and noise.      
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial markets span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd''. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to create trends in asset prices. 
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based.

Towards the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery whereby information is assimilated into price.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to the movement of prices in the foreign exchange market\footnote{The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}.  If speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

This paper contributes to the debate about the nature of speculation by providing clear evidence that speculation is more informed than uninformed. A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to assess the price action around extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.

@@

\section{Informed and uninformed speculation}
Speculation plays an important part in several areas of financial economics.  However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making. A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.   Speculation and trading can be categorised as being informed or uninformed. Those trading without information are frequently called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.

The distinction between informed and uninformed traders is a central component of microstructure research.  The focus of the \citet{Kyle1985Continuous} model of market-making is the interaction of orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.  In a similar way \citet{glosten1985bid} describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.

\citet[p.529]{BlackNoise} suggested that the interaction of informed and uninformed activity was essential to the smooth functioning of financial markets. Noise, he argued, was the  ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

@@

\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.  Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{Market-makers use Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

@@

Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.  Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.   This leaves uninformed speculation as a pure noise that can be characterised as white-noise random variable.\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.} While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

Noise-trading caused by psychological biases and restricted information is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred (perhaps incorrectly) from the buying and selling conducted by other traders, causing positive feedback and trends.\footnote{This is an example of the beauty contest that has been presented by \citet[p.101]{Keynes1936}.}

Liquidity trading suggests that institutional constraints may ensure that economic agents just execute customer orders at the current price or that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are also associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.

@@

  The De-Long noise-trader model also provides a useful benchmark against which to assess the activity of speculators. The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, is the original arbitrage from the misperception of the noise-traders; \emph{Price Pressure} is the \emph{noise-trader risk} that prices may diverge further before returning to equilibrium; \emph{Hold More} means that noise-traders become more involved in the market as their views become more extreme; \emph{Create Space} is the difference between noise-trader perception and the actual value, it creates the opportunity for informed traders to profit from their information. The aim of this paper is to establish whether the Friedman-effect can be identified.

\section{Measuring speculation and the event study}
There are two steps: identify speculation and assess what happens when it becomes extreme.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated; if informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium and the following price performance should be a random walk.

The \citet[pp. 14-15]{Delong1990noise} presented a widely-used model to understand the interaction of informed and uninformed traders.  It is summarised by Equation \ref{eq:NP2}. There are two parameters that determine the size of the reversal or Friedman-effect that is to be expected:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If speculation is uninformed and either of these parameters reach an extreme, prices should subsequently reverse. The following paragraphs show how these can be measured. \footnote{See \citet{Hayward2013} for full details of a Monte-Carlo simulations of this model when the speculative misperception $(\rho)$ and the weight of speculation $(\mu)$ parameters are adjusted. As the expected deviation from fundamental value and the weight of speculators in the market becomes more extreme, this reversal becomes more likely.} This reversal is the Friedman-effect discussed above.

The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the absolute value of $\rho - 1$.  In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.} The risk-reversal skew is the relationship between a call option and the equivalent put. The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

@@

Recording speculative activity in an over-the-counter market like foreign exchange is extremely difficult.  However, there are some exchange-traded currency derivative markets in the US, and participants in these markets are required to report their positions to the US regulators.  The Commodity Futures Trading Commission (CFTC) collect weekly information about the open positions held by private entities in the US derivative markets.   A sub-set of this information is released to the public each week as the \emph{Commitment of Traders Report} (CoT).\footnote{The \citet{cot} require traders to categorise themselves as being either commercial (C) or non-commercial (NC).  Commercial traders must have some underlying business interest in the security, commodity or instrument that they are trading.  These could be seen as natural hedgers.  Non-commercial accounts are generally considered to be speculators.  See \citet{FuturesSanders} and \citet{FuturesWang} for examples of previous use of this data in this fashion.

The data used run from September 30 1998 when the data started to be released on a weekly basis and continue up to December 31 2008.  The Euro data runs from March 3 1998 to December 31 2008.  There are two gaps in the Swiss Franc data for September 14 2004 and September 21 2004.  The absence of these data seems to be a result of problems with the CFTC database.  The contracts are for EUR 100,000; GBP 62,500; JPY 12,500,000; CHF 100,000 and CAD 100,000.}  
An \emph{event study} is used to assess the effect of extreme speculation on future returns.  See \citet{Dolly1933} for the original version of the event study method and \citet{FamaFisherJensenRoll} for an overview of the method used in this study. The R package \emph{eventstudies} created by \citet{eventstudies} is used to manage the data.  Extreme measures of speculative sentiment and activity are the event and the cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) are the exchange rates to be assessed.

\section{Analysis of results}
  Table \ref{tabref:RR1} records the result of the event studies that were carried out on extreme risk reversal.   For each exchange rate the second column distinguishes between extreme high and extreme low, the third  column identifies the number of extreme events that were recorded and the following two are the cumulative average returns for the whole event window (CARW) and the cumulative average returns for the event and the period after (CARA).  The windows were chosen to correspond to one day and one week of trading.  However, Figure \ref{fig:ES1}, which shows the extreme high for a 16 day event window, makes it clear that altering the size of the event window does not change the findings to any significant degree.
\begin{sidewaystable}
% multirow package is required.   
%pdflandscape is required to facilitate the landscape position. 
\begin{threeparttable}
\caption{Event Study: Cumulative Abnormal Returns and Extreme Risk Reversal}
\begin{tabular}{llccccccccccccc}  
 % \hline
& &\multicolumn{6}{c}{$5^{th}$ and $95^{th}$ percentile} & \multicolumn{1}{c}{} & \multicolumn{6}{c}{$1^{st}$ and $99^{th}$ percentile}\\
 & &\multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} & \multicolumn{1}{c}{} & \multicolumn{3}{c}{1 day} & \multicolumn{3}{c}{4 day} \\ 
%\cline{4 - 10} \cline{13 - 19}
& &\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA} & \multicolumn{1}{c}{} & 
\multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  
& \multicolumn{1}{ c}{N} & \multicolumn{1}{c}{CARW} & \multicolumn{1}{c}{CARA}  \\   
\hline
\multirow{2}{*}{EUR} 
& Hi &  87 &  0.4103* &0.1586 &87 &0.9442* & 0.1505 &&22 &0.0309 & -0.1748 &22 & 0.5781*& -0.5387*  \\ 
& Lo & 106 &-0.3313* & -0.0933 &106 &-0.0816* &-0.1991 & & 21&0.3040 &0.1820 &21 &0.7935* & 0.4377  \\
\multirow{2}{*}{JPY}
& Hi & 90 &0.4860* &0.1341 &90 &1.5158* &0.2531 & &20 &0.8227* &0.3078* &20 &2.4337* & 0.8284*  \\ 
& Lo & 90 &-0.3168 &0.0945 &90 &-1.3532* &0.1984 & &17 &-1.6990* &-0.9347* &17 &-3.6751* & -1.2556*  \\
\multirow{2}{*}{GBP}
& Hi & 96 &0.3303* &0.1903* &96 &0.8966* &0.2547 & &21 &0.5252* &0.1774 &21 &1.3131* &0.3284  \\ 
& Lo & 97 &-0.1894 &0.0312 &97 &-0.8031* &0.1500 & &24 &-0.2630 &-0.0866 &24 &-0.6660 &0.1888  \\
\multirow{2}{*}{CHF}
& Hi & 80 &0.0065 &-0.0565 &80 &0.5624* &0.1322 & &18 & -0.0022 & -0.1518 &18 &0.4887 &-0.1768  \\ 
& Lo & 75 &-0.3716* & -0.1125& 75 &-1.0071* &-0.0219 & &18 &0.0064 &0.1704 &18 &-0.8661* &0.7631  \\
\multirow{2}{*}{CAD}
& Hi & 113 &-0.0038 &-0.0489 &113 &0.1872 &-0.1951 & & 24 &-0.1081 & -0.0462 &24 &0.3935* & -0.1592  \\ 
